Magnetic grid resolution enhancement using machine learning: A case study from the Eastern Goldfields Superterrane

### 2023-07-20 paper1.tex
Include figures and tables

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/fig1.jpg}
    \caption[Comparison of different resolution grids]{Low resolution \SI{80}{\metre} cell size (a), and high-resolution \SI{20}{\metre} cell size (b) TMI grids over the same extent.
    Each tile is \SI{2560}{\metre} \texttimes\ \SI{2560}{\metre}.
    The additional detail shown in the high-resolution grid is made available by the finer cell size, which can convey higher frequency detail, if present in the sampled data.}
    \label{lrandhr}
\end{figure}

@@

Neural network based super-resolution has seen recent success across many scientific disciplines on data other than natural images.
It is popular in satellite remote sensing \parencite[e.g.][]{lanarasSuperresolutionSentinel2Images2018b,wangUltradenseGANSatellite2019}, and has been applied to astrophysics \parencite[e.g.][]{jungbluthSingleframeSuperresolutionSolar2019,salvatelliUsingUNetsCreate2019}, fluid dynamics \parencite{bodeUsingPhysicsinformedEnhanced2021}, elevation models \parencite{leongDeepBedMapDeepNeural2020}, hyperspectral data \parencite[e.g.][]{arunCNNBasedSuperResolutionHyperspectral2020}, and seismic geophysics data \parencite[e.g.][]{liDeepLearningSimultaneous2021,wangDeeplearningbasedSeismicData2018,wangAdaptingResidualDense2022,yuDeepLearningDenoising2019}.
These authors typically report a lower error between the super-resolution upsampled product and ground truth data when compared to a bicubic upsampled product and the ground truth data, as well as improved perceptual quality.
Critically, super-resolution for these applications was able to create realistic high frequency information during upsampling, which could be validated against existing high-resolution surveys of the same extent.
While there are pre-trained models to super-resolve image data (such as ImageNet), the smooth and continuously differentiable nature of potential field data makes it difficult to directly apply such pre-trained models.
For example, it may enhance the appearance of gridding artifacts, instead of supressing them.
While several of the aforementioned works investigate deep learning super-resolution for geospatial data, low resolution in potential field geophysics is a result of under sampling and naïve best-effort interpolation of the potential field, distinct from low-resolution data previously studied in those works.

\begin{figure}
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[The RDN architecture]{The RDN architecture, which acts as the ESRGAN generator.
    Twenty-three Basic Blocks are stacked sequentially to form the main convolution path.
    Upsampling to the target output size is performed after the main convolution sequence.
    Adapted from \cite{wangESRGANEnhancedSuperResolution2018}.
    }
    \label{fig:rdn}
\end{figure}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[Basic Blocks]{Each Basic Block (\cref{fig:rdn}, orange) comprises three weighted Residual Dense Blocks (\cref{fig:convs}).
    ESRGAN+ extends ESRGAN with the addition of Gaussian Noise after each basic RDB block (black vectors).
    Weight factors are indicated by $\beta$ and $\gamma$.
    Noise injection is not used in RDN\@.
    Adapted from \cite{rakotonirinaESRGANFurtherImproving2020}.
    }
    \label{fig:rdb}
\end{figure}

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[RDB convolutional layers]{Each Residual Dense Block (\cref{fig:rdb}, purple) comprises four convolutional layers, with additional learning pathways present as residual connections (black), and dense connections (green).
    LReLU refers to the Leaky-ReLU activation function.
    Adapted from \cite{rakotonirinaESRGANFurtherImproving2020}.
    }
    \label{fig:convs}
\end{figure}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[Dataset context]{Reserved training/validation and test areas of Eastern Goldfields Superterrane magnetic grid, and their tectonic context.
    Background: \SI{20}{\metre} merged magnetic grid of Western Australia, \parencite{brett40MagneticMerged2020}.
}
    \label{fig:trainvaldata}
\end{figure}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[Visual inference results]{Visual comparisons of a selected tile from the reserved test region.
    The \SI{80}{\metre} low-resolution input (a) is shown at equivalent size to the \SI{20}{\metre} high-resolution ground truth (b) using nearest neighbour interpolation.
    The colourmap is shared between subfigures.
    }
    \label{fig:resultsvis}
\end{figure}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[Error residuals map for a selected tile]{Error residuals map for the selected test tile, for each upsampling method; Bicubic (a), RDN\textdaggerdbl{} (b), and ESRGAN+ (c).
    Image quality metrics are reported for each upsampled tile, calculated against the groundtruth HR\@.}
    \label{fig:residuals}
\end{figure}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[short]{Radially Averaged Power Spectra for the complete test extent shown in \cref{fig:trainvaldata}.
    The Nyquist wavelengths for the high-resolution (and upsampled) rasters and low-resolution raster are also shown.
    The frequency axis has been transformed to wavelength, and the spectra have been normalised for direct comparison.
    Note the complete test extent is covered by surveys with a range of line spacings from \SI{100}{\metre} to \SI{20}{\metre}, with \SI{100}{\metre} being the most extensive.
    The Nyquist wavelength of the \SI{100}{\metre} surveys (optimal grid cell size approximately \SI{20}{\metre}) likely causes the reduction in power at \SI{55}{\metre} and \SI{220}{\metre}.
    }
    \label{fig:rapsd}
\end{figure}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[High Pass filtered results]{High Pass filtered test tile and its upsampled counterparts.
    The cut-off for the high pass Butterworth filter was selected with wavelength equal to the Nyquist wavelength of the LR grid, i.e. four times the HR Nyquist wavelength
    }
    \label{fig:filtered}
\end{figure}

\begin{table}
    \begin{tabularray}
    \end{tabularray}
    % \begin{tabularray}{l|ll|ll|ll}
    %             & Bicubic        &                          & RDN\textdaggerdbl{}     &                          & ESRGAN+       &                    \\
    %             & Mean           & Standard Deviation       & Mean                    & StandardDeviation       & Mean          & Standard Deviation \\
    %     \hline{}
    %     RMSE       & \num{13.545}   & \num{26.576}             & \num{7.265}             & \num{13.741}             & \num{11.120}  & \num{21.670}             \\
    %     SISM       &  \num{0.924}   &  \num{0.047}             & \num{0.959}             &  \num{0.037}             &  \num{0.922}  &  \num{0.054}             \\
    %     MS-SSIM    &  \num{0.971}   &  \num{0.018}             & \num{0.988}             &  \num{0.012}             &  \num{0.976}  &  \num{0.020}             \\
    %     \noalign{}

    % \end{tabularray}
    \caption[Accuracy Metrics]{
        Accuracy metrics for each upsampling method.
        The best performing method for each metric is bolded.
     }
    \label{tab:metrics}
\end{table}

@@

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[First Vertical Derivative SR TMI over the Mt Mason resource project]{First Vertical Derivative TMI over the Mt Mason resource project, with the super-resolution predictions (d, e) showing a variety of inference artefacts.
    Red asterisks (*) show a vertical tile boundary present as a 3--4 pixel-wide error in predicted TMI value.
    Red caret (\^{ }) indicates a region of subtle ripple-mark textural artefact, only present in ESRGAN+ predictions.
    }
    \label{fig:jhillvis}
\end{figure}

\begin{figure}[hbt]
    % \includegraphics[width=\textwidth]{fig/p1/}
    \caption[Jasper Hills Project TMI grids and line profile]{Jasper Hills Project TMI grids (a--e) and line profile (f).
    Bicubic interpolation of the low-resolution grid closely matches the low-resolution grid, while both super-resolution methods better approximate the magnitude of the features present in the HR profile, between \SI{650}{\metre} and \SI{1020}{\metre}.
    }
    \label{fig:jhilldykes}
\end{figure}

>>>>

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/image1.png}
    \caption[Comparison of different resolution grids]{Low resolution \SI{80}{\metre} cell size (a), and high-resolution \SI{20}{\metre} cell size (b) TMI grids over the same extent.
    Each tile is \SI{2560}{\metre} \texttimes\ \SI{2560}{\metre}.
    The additional detail shown in the high-resolution grid is made available by the finer cell size, which can convey higher frequency detail, if present in the sampled data.}
    \label{lrandhr}
\end{figure}

@@

Neural network based super-resolution has seen recent success across many scientific disciplines on data other than natural images.
It is popular in satellite remote sensing \parencite[e.g.][]{lanarasSuperresolutionSentinel2Images2018,wangUltradenseGANSatellite2019}, and has been applied to astrophysics \parencite[e.g.][]{jungbluthSingleframeSuperresolutionSolar2019,salvatelliUsingUNetsCreate2019}, fluid dynamics \parencite{bodeUsingPhysicsinformedEnhanced2021}, elevation models \parencite{leongDeepBedMapDeepNeural2020}, hyperspectral data \parencite[e.g.][]{arunCNNBasedSuperResolutionHyperspectral2020}, and seismic geophysics data \parencite[e.g.][]{liDeepLearningSimultaneous2021,wangDeeplearningbasedSeismicData2018,wangAdaptingResidualDense2022,yuDeepLearningDenoising2019}.
These authors typically report a lower error between the super-resolution upsampled product and ground truth data when compared to a bicubic upsampled product and the ground truth data, as well as improved perceptual quality.
Critically, super-resolution for these applications was able to create realistic high frequency information during upsampling, which could be validated against existing high-resolution surveys of the same extent.
While there are pre-trained models to super-resolve image data (such as ImageNet), the smooth and continuously differentiable nature of potential field data makes it difficult to directly apply such pre-trained models.
For example, it may enhance the appearance of gridding artifacts, instead of supressing them.
While several of the aforementioned works investigate deep learning super-resolution for geospatial data, low resolution in potential field geophysics is a result of under sampling and naïve best-effort interpolation of the potential field, distinct from low-resolution data previously studied in those works.

\begin{figure}
    \includegraphics[width=\textwidth]{fig/p1/image2.png}
    \caption[The RDN architecture]{The RDN architecture, which acts as the ESRGAN generator.
    Twenty-three Basic Blocks are stacked sequentially to form the main convolution path.
    Upsampling to the target output size is performed after the main convolution sequence.
    Adapted from \cite{wangESRGANEnhancedSuperResolution2018}.
    }
    \label{fig:rdn}
\end{figure}

@@

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/image3.png}
    \caption[Basic Blocks]{Each Basic Block (\cref{fig:rdn}, orange) comprises three weighted Residual Dense Blocks (\cref{fig:convs}).
    ESRGAN+ extends ESRGAN with the addition of Gaussian Noise after each basic RDB block (black vectors).
    Weight factors are indicated by $\beta$ and $\gamma$.
    Noise injection is not used in RDN\@.
    Adapted from \cite{rakotonirinaESRGANFurtherImproving2020}.
    }
    \label{fig:rdb}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/image4.png}
    \caption[RDB convolutional layers]{Each Residual Dense Block (\cref{fig:rdb}, purple) comprises four convolutional layers, with additional learning pathways present as residual connections (black), and dense connections (green).
    LReLU refers to the Leaky-ReLU activation function.
    Adapted from \cite{rakotonirinaESRGANFurtherImproving2020}.
    }
    \label{fig:convs}
\end{figure}

@@

\begin{figure}[]
    \includegraphics[width=\textwidth]{fig/p1/image5.png}
    \caption[Dataset context]{Reserved training/validation and test areas of Eastern Goldfields Superterrane magnetic grid, and their tectonic context.
    Background: \SI{20}{\metre} merged magnetic grid of Western Australia, \parencite{brett40MagneticMerged2020}.
}
    \label{fig:trainvaldata}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\textwidth]{fig/p1/image6.png}
    \caption[Visual inference results]{Visual comparisons of a selected tile from the reserved test region.
    The \SI{80}{\metre} low-resolution input (a) is shown at equivalent size to the \SI{20}{\metre} high-resolution ground truth (b) using nearest neighbour interpolation.
    The colourmap is shared between subfigures.
    }
    \label{fig:resultsvis}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\textwidth]{fig/p1/image7.png}
    \caption[Error residuals map for a selected tile]{Error residuals map for the selected test tile, for each upsampling method; Bicubic (a), RDN\textdaggerdbl{} (b), and ESRGAN+ (c).
    Image quality metrics are reported for each upsampled tile, calculated against the groundtruth HR\@.}
    \label{fig:residuals}
\end{figure}

@@

% TODO USE REVISED MANUSCRIPT FIGURES

\begin{figure}
    \includegraphics[width=\textwidth]{fig/p1/image8.png}
    \caption[short]{Radially Averaged Power Spectra for the complete test extent shown in \cref{fig:trainvaldata}.
    The Nyquist wavelengths for the high-resolution (and upsampled) rasters and low-resolution raster are also shown.
    The frequency axis has been transformed to wavelength, and the spectra have been normalised for direct comparison.
    Note the complete test extent is covered by surveys with a range of line spacings from \SI{100}{\metre} to \SI{20}{\metre}, with \SI{100}{\metre} being the most extensive.
    The Nyquist wavelength of the \SI{100}{\metre} surveys (optimal grid cell size approximately \SI{20}{\metre}) likely causes the reduction in power at \SI{55}{\metre} and \SI{220}{\metre}.
    }
    \label{fig:rapsd}
\end{figure}

@@

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/image9.png}
    \caption[High Pass filtered results]{High Pass filtered test tile and its upsampled counterparts.
    The cut-off for the high pass Butterworth filter was selected with wavelength equal to the Nyquist wavelength of the LR grid, i.e. four times the HR Nyquist wavelength
    }
    \label{fig:filtered}
\end{figure}

\begin{table}
    \begin{tblr}{
        colspec={
            r
            Q[si={table-format=2.3},r]
            Q[si={table-format=2.3},l]
            Q[si={table-format=2.3},r]
            Q[si={table-format=2.3},l]
            Q[si={table-format=2.3},r]
            Q[si={table-format=2.3},l]
            },
        row{1-2} = {guard}
    }
    & Bicubic & & RDN\textdaggerdbl{} & & ESRGAN+ & \\
    \hline{}
    & Mean & {Standard \\ Deviation} & Mean & {Standard \\ Deviation} & Mean & {Standard \\ Deviation} \\
    RMSE   & 13.545 & 26.576 & 7.265 & 13.741 & 11.120 & 21.670 \\
    SISM    & 0.924 &  0.047 & 0.959 &  0.037 &  0.922 &  0.054 \\
    MS-SSIM & 0.971 &  0.018 & 0.988 &  0.012 &  0.976 &  0.020 \\
    \end{tblr}

    \caption[Accuracy Metrics]{
        Accuracy metrics for each upsampling method.
        The best performing method for each metric is bolded.
     }
    \label{tab:metrics}
\end{table}

@@

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/image10.png}
    \caption[First Vertical Derivative SR TMI over the Mt Mason resource project]{First Vertical Derivative TMI over the Mt Mason resource project, with the super-resolution predictions (d, e) showing a variety of inference artefacts.
    Red asterisks (*) show a vertical tile boundary present as a 3--4 pixel-wide error in predicted TMI value.
    Red caret (\^{ }) indicates a region of subtle ripple-mark textural artefact, only present in ESRGAN+ predictions.
    }
    \label{fig:jhillvis}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\textwidth]{fig/p1/image11.png}
    \includegraphics[width=\textwidth]{fig/p1/image12.png}
    \caption[Jasper Hills Project TMI grids and line profile]{Jasper Hills Project TMI grids (a--e) and line profile (f).
    Bicubic interpolation of the low-resolution grid closely matches the low-resolution grid, while both super-resolution methods better approximate the magnitude of the features present in the HR profile, between \SI{650}{\metre} and \SI{1020}{\metre}.
    }
    \label{fig:jhilldykes}
\end{figure}
